---
title: "R Notebook"
output: html_notebook
---

Data Frames
The function data.frame() creates data frames, tightly coupled collections of variables which share many of the properties of
matrices and of lists, used as the fundamental data structure by most of R's modeling software.
```{r}
mtcars
tail(mtcars)

str(mtcars)
```

```{r}
name <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)

planets_df<-data.frame(name,type,diameter,rotation,rings)
```



Selection of data frame elements
```{r}
planets_df[1,3]
planets_df[4,]
planets_df[1:5,]
planets_df[1:5,c("rotation", "type")]

rings_vector<- planets_df$rings
rings_vector
subset()

```
Sorting:
In data analysis you can sort your data according to a certain variable in the data set. In R, this is done with the help of the function order().

Order() is a function that gives you the ranked position of each element when it is applied on a variable, such as a vector for example:
```{r}
a<-c(100,10,1000)
order(a)
b<-a[order(a)]

positions<-order(planets_df$diameter)
planets_df[positions,]  #para seleccionar datos en un dataframe, se pone (fila, columna) y cuando mando a llamar solo uno de los dos, asi obtengo todos los datos de esa posicion
```

Lists
A list in R is similar t your to-do lista at work or school: the different items on that list most likely differ in lenght, characteristic, and type of activity that has to be done.
A list in R allows you to gather a variety of objects under one name (that is the name of the list) in an ordered way. These objects can be matrices, vectors, data frames, even other lists, etc. It is not even required that these objects are related to each other in any way.
You could say that a list is some kind super data type: you can store practically any piece of information in it!

```{r}
my_vector<-1:10
my_matrix<-matrix(1:9,ncol=3)
my_df<-planets_df[1:4,]

my_list<-list(my_vector,my_matrix,my_df)
```

Creating a named list

```{r}
names(my_list)<-c("vector","matrix","data frame")
my_list
```
Selecting elements from a list
One way to select a component is using the numbered position of that component
```{r}
my_list[[2]][2,]
my_list[[3]][2:3,c("name", "diameter")]
```


