---
title: "R Notebook"
output: html_notebook
---

Grouping and summarizing:
The summarize verb-- mean() and sum() are only two of the funtions R provides for summarizing a collection of numbers. 

```{r}
library(gapminder)
library(dplyr)
library(tidyverse)

gapminder %>%
  summarize(medianLifeExp=median(lifeExp))

gapminder %>%
  filter(year==1957) %>%
  summarize(medianLifeExp=median(lifeExp))

gapminder %>%
  filter(year==1957) %>%
  summarise(medianLifeExp=median(lifeExp), max(gdpPercap))
```
The group_by verb
Most data operations are done on groups defined by variables. group_ny() takes an existing data frame and converts it into a grouped data frame where operations are performed "by group"
```{r}
gapminder %>%
  group_by(year) %>%
  summarise(medianLifeExp = median(lifeExp), max(gdpPercap))
```
```{r}
gapminder %>%
  group_by(continent, year) %>%
  summarise(medianLifeExp = median(lifeExp), max(gdpPercap))
```
```{r}
by_year <-gapminder %>%
  group_by(year) %>%
  summarise(medianLifeExp = median(lifeExp), max(gdpPercap))

ggplot(by_year, aes(x=year, y=medianLifeExp))+
  geom_point()
```
```{r}
by_year_continent <- gapminder %>%
  group_by(continent, year) %>%
  summarise(medianGdpPercap = median(gdpPercap))

ggplot(by_year_continent, aes(x=year, y=medianGdpPercap, color=continent))+
  geom_point()
```
Line Plots:
A line plot is useful for visualizing trends over time. In this excercise, you'll examine how the median GDP per capita has changed over time.
```{r}
by_year<- gapminder %>%
  group_by(year) %>%
  summarise(medianGdpPercap = median(gdpPercap))

ggplot(by_year, aes(x=year, y=medianGdpPercap))+
  geom_line()
```
```{r}
by_year_continent<- gapminder %>%
  group_by(year, continent) %>%
  summarise(medianGdpPercap = median(gdpPercap))

ggplot(by_year_continent, aes(x=year, y=medianGdpPercap, color=continent))+
  geom_line()

```
Bar Plots:
A bar plot is useful for visualizing summary statistics, such as the median GDP in each continent
```{r}
oceania_1952 <- gapminder %>%
  filter(continent == "Oceania", year==1952)

oceania_1952

ggplot(oceania_1952, aes(x=country, y=gdpPercap, fill=country))+
  geom_col()
```

Histograms:
A histogram is useful for examining the distribution of a numeric variable.
```{r}

gapminder_1952<- gapminder%>%
  filter(year==1952) %>%
  mutate(pop_by_mil=pop/1000000)

ggplot(gapminder_1952, aes(x=pop_by_mil))+
  geom_histogram(bins = 50)

#los bins son la cantidad de rangos que se establecen dentro de la grafica.
```
```{r}
gapminder_1952<- gapminder%>%
  filter(year==1952) %>%
  mutate(pop_by_mil=pop/1000000)

ggplot(gapminder_1952, aes(x=pop_by_mil))+
  geom_histogram(bins = 50)+
  scale_x_log10()
```
Box Plots
A boxplot is useful for comparing a distribution of values across several groups, in this excercise youll examine the distribution of GDP per capita by continent.

```{r}

gapminder_1952<- gapminder%>%
  filter(year==1952)

ggplot(gapminder_1952, aes(x=continent, y=gdpPercap, fill=continent))+
  geom_boxplot()+
  scale_y_log10()
```

Data manipulation with dplyr
```{r}
counties<- read.csv("counties.csv")
```
```{r}
counties %>%
  select(state, county, population, men, women) %>%
  mutate(proportion_men = men/population) %>%
  filter(population >1000) %>%
  arrange(desc(proportion_men))
  
```

